---
title: "Figure6"
author: "Rahul Samrat"
date: "2025-03-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Figure 6a. PCA showing lipid profile different in Climgrass treatments

```{r}
# Load necessary libraries
library(readr)
library(dplyr)
library(ggplot2)
library(FactoMineR)
library(vegan)
library(ggrepel)
library(tibble)
```

```{r}
# Load the data
metadata <- read_csv('Metadata.csv')
normalized_data <- read_csv('normalized_data.csv')

# Standardize sample names to avoid mismatches
metadata$Sample <- as.character(trimws(metadata$Sample))
colnames(normalized_data) <- as.character(trimws(colnames(normalized_data)))

# Transpose normalized data and merge with metadata
normalized_data_t <- as.data.frame(t(normalized_data[-1]))
colnames(normalized_data_t) <- normalized_data$GlobalID
rownames(normalized_data_t) <- colnames(normalized_data)[-1]

# Prepare metadata for merging
metadata <- metadata %>% column_to_rownames(var = "Sample")

# Merge data on sample names
merged_data <- merge(metadata, normalized_data_t, by = "row.names")
rownames(merged_data) <- merged_data$Row.names
merged_data <- merged_data[-1]

# Extract dependent and independent variables
independent_vars <- merged_data[, 1:2]
dependent_vars <- merged_data[, -c(1:2)]

# Check for negative values in the data
if (any(dependent_vars < 0)) {
  print("Negative values found in dependent variables. Correcting data...")
  dependent_vars[dependent_vars < 0] <- 0
}

```

# Recompute the distance matrix with Bray-Curtis method

```{r}
distance_matrix <- vegdist(dependent_vars, method = "bray")
```

# Perform PCA

```{r}
pca_result <- PCA(dependent_vars, graph = FALSE)
pca_data <- as.data.frame(pca_result$ind$coord[, 1:2])
colnames(pca_data) <- c("PC1", "PC2")
pca_data <- cbind(pca_data, independent_vars)

# Create a combined group factor
pca_data$Combined_Group <- ifelse(pca_data$Type2 == "Drought", "Drought", "No_Drought")

# Compute Bray-Curtis distance matrix
distance_matrix <- vegdist(dependent_vars, method = "bray")
```

# Perform PERMANOVA

```{r}
result <- adonis2(distance_matrix ~ Combined_Group, data = pca_data, method = "bray", permutations = 999)

permanova_result <- adonis2(distance_matrix ~ Combined_Group, data = pca_data, permutations = 9999)
permanova_p <- permanova_result$`Pr(>F)`[1]
```

# Plot PCA with convex hulls and PERMANOVA annotation

```{r}
# Define color and shape mappings
colors <- c("Ambient" = "#0072B2", "Future" = "#D55E00")
shapes <- c("Drought" = 16, "No_drought" = 17)
hull_colors <- c("Drought" = "#E69F00", "No_Drought" = "#56B4E9")


pca_plot <- ggplot(pca_data, aes(x = PC1, y = PC2, color = Type1, shape = Type2)) +
  geom_point(size = 4, alpha = 0.8) +
  scale_color_manual(values = colors, name = "Climate Scenario") +
  scale_shape_manual(values = shapes, name = "Water Condition") +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "gray90"),
    legend.position = "right",
    legend.box = "vertical",
    legend.key.size = unit(1, "cm"),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 14, face = "bold"),
    plot.title = element_text(hjust = 0.5, face = "bold", size = 16),
    axis.title = element_text(face = "bold", size = 14),
    axis.text = element_text(size = 12)
  ) +
  labs(
    title = "PCA of Soil Lipid Profiles",
    x = paste0("PC1 (", round(pca_result$eig[1, 2], 1), "%)"),
    y = paste0("PC2 (", round(pca_result$eig[2, 2], 1), "%)")
  ) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  coord_fixed(ratio = 1)

# Add convex hulls
for (group in unique(pca_data$Combined_Group)) {
  group_data <- pca_data[pca_data$Combined_Group == group, ]
  if (nrow(group_data) > 2) {
    hull_indices <- chull(group_data$PC1, group_data$PC2)
    hull_data <- group_data[hull_indices, ]
    pca_plot <- pca_plot + 
      geom_polygon(data = hull_data, aes(x = PC1, y = PC2, fill = Combined_Group),
                   alpha = 0.2, color = hull_colors[group], show.legend = FALSE)
  }
}

# Add PERMANOVA annotation
pca_plot <- pca_plot + 
  annotate("text", x = Inf, y = Inf, label = paste0("PERMANOVA\np = ", signif(permanova_p, 2)), 
           hjust = 1.1, vjust = 1.1, size = 5, 
           fontface = "bold", color = "black", 
           box.padding = unit(0.35, "lines"))

# Save and display the plot
ggsave("publication_quality_pca_plot.pdf", plot = pca_plot, width = 10, height = 8, dpi = 300)
print(pca_plot)
```

# Figure 6b. Shannon entropy (lipid diversity) across climgrass treatments

## Calculating concentration of each lipid with different treatment

```{r}
# Load necessary libraries
library(readxl)
library(dplyr)
library(tidyr)
library(ggplot2)
library(viridis)  # For a colorblind-friendly palette
library(ggthemes) # For additional themes
library(extrafont) # For using custom fonts (you may need to install this)
library(readxl)
library(dplyr)
library(writexl)
library(tidyverse)
library(ggrepel)
library(scales)
library(gridExtra)


# Load the Excel file
file_path <- 'Fig5-ALL-concentration-updated.xlsx'
df <- read_excel(file_path, sheet = 'SIRIUS-STD')

# Sample data with combined treatment type
sample_data <- data.frame(
  Sample = c(19, 22, 36, 28, 44, 49, 35, 40, 52, 27, 43, 48, 50),
  Type1 = c("Ambient", "Ambient", "Ambient", "Ambient", "Ambient", "Ambient", "Future", "Future", "Future", "Future", "Future", "Future", "Future"),
  Type2 = c("No_drought", "No_drought", "No_drought", "Drought", "Drought", "Drought", "No_drought", "No_drought", "No_drought", "Drought", "Drought", "Drought", "Drought"),
  Weight = c(2.054, 1.965, 2.063, 2.048, 2.063, 2.082, 2.065, 2.032, 2.033, 2.071, 2.028, 2.027, 0.178)
)
sample_data$Combined_Treatment <- paste(sample_data$Type1, sample_data$Type2, sep = "_")

# Mapping sample numbers to treatment columns
sample_to_column <- list(
  `19` = 'Ambient_No_drought_Sample_19',
  `22` = 'Ambient_No_drought_Sample_22',
  `36` = 'Ambient_No_drought_Sample_36',
  `28` = 'Ambient_Drought_Sample_28',
  `44` = 'Ambient_Drought_Sample_44',
  `49` = 'Ambient_Drought_Sample_49',
  `35` = 'Future_No_drought_Sample_35',
  `40` = 'Future_No_drought_Sample_40',
  `52` = 'Future_No_drought_Sample_52',
  `27` = 'Future_Drought_Sample_27',
  `43` = 'Future_Drought_Sample_43',
  `48` = 'Future_Drought_Sample_48'
)

# Filter out rows where the corresponding LOQ column value is not 1 (good value)
concentration_cols <- grep('Concentration$', names(df), value = TRUE)
loq_cols <- grep('Concentration > LOQ$', names(df), value = TRUE)
for (loq_col in loq_cols) {
  df <- df %>% filter(df[[loq_col]] == 1)
}

# Creating a new dataframe to hold the rearranged columns
filtered_df <- df %>% select(GlobalID, `Molecule List`, LipidMolecule)

# Renaming the columns based on combined treatment type and sample number
for (i in 1:nrow(sample_data)) {
  sample <- sample_data$Sample[i]
  treatment_col <- paste(sample_data$Combined_Treatment[i], 'Sample', sample, sep = "_")
  sample_col <- paste(sample, 'Sum Total Area Concentration', sep = " ")
  if (sample_col %in% names(df)) {
    filtered_df[[treatment_col]] <- df[[sample_col]]
  }
}

# Creating the new dataframe with the desired format including GlobalID
new_format_data_with_id <- data.frame(
  GlobalID = integer(),
  Molecule.List = character(),
  LipidMolecule = character(),
  Sample = integer(),
  Concentration = numeric(),
  Treatment = character(),
  stringsAsFactors = FALSE
)

# Populating the new dataframe with GlobalID
for (i in 1:length(sample_to_column)) {
  sample <- as.numeric(names(sample_to_column)[i])
  column <- sample_to_column[[i]]
  treatment_type <- strsplit(column, '_Sample_')[[1]][1]
  for (j in 1:nrow(filtered_df)) {
    new_format_data_with_id <- rbind(new_format_data_with_id, data.frame(
      GlobalID = filtered_df$GlobalID[j],
      Molecule.List = filtered_df$`Molecule List`[j],
      LipidMolecule = filtered_df$LipidMolecule[j],
      Sample = sample,
      Concentration = filtered_df[[column]][j],
      Treatment = treatment_type
    ))
  }
}

output_path <- "information-theory-indices1.xlsx"
write_xlsx(new_format_data_with_id, output_path)

# Display the final dataframe
print(head(new_format_data_with_id))
```

# Calculation of Shannon entropy Hj

```{r}
# Load necessary libraries
library(readxl)
library(dplyr)
library(ggplot2)

# Load data
file_path <- "information-theory-indices1.xlsx"
data <- read_excel(file_path, sheet = "Sheet1")

# Calculate relative frequency (P_ij)
data <- data %>%
  group_by(Sample) %>%
  mutate(P_ij = Concentration / sum(Concentration)) %>%
  ungroup()

# Calculate Shannon entropy (H_j) per sample
H_j <- data %>%
  group_by(Sample) %>%
  summarise(H_j = -sum(P_ij * log2(P_ij), na.rm = TRUE))

# Merge with treatment information explicitly
sample_treatments <- data %>% select(Sample, Treatment) %>% distinct()
H_j <- left_join(H_j, sample_treatments, by = "Sample")

# Ensure factor levels match the actual treatment labels
H_j$Treatment <- factor(H_j$Treatment, 
                        levels = c("Ambient_No_drought", "Ambient_Drought", "Future_No_drought", "Future_Drought"))

# Perform one-way ANOVA safely after ensuring factors
anova_result <- aov(H_j ~ Treatment, data = H_j)
anova_summary <- summary(anova_result)
anova_pvalue <- anova_summary[[1]][["Pr(>F)"]][1]

# Define custom colors for treatments explicitly matching actual labels
treatment_colors <- c("Ambient_No_drought" = "#A6CEE3",
                      "Ambient_Drought" = "#FDBF6F",
                      "Future_No_drought" = "#1F78B4",
                      "Future_Drought" = "#E31A1C")

# Plot with annotation
shannon_plot <- ggplot(H_j, aes(x = Treatment, y = H_j, fill = Treatment)) +
  geom_boxplot(alpha = 0.7, color = "black") +
  geom_jitter(width = 0.15, size = 2, alpha = 0.6, color = "black") +
  scale_fill_manual(values = treatment_colors) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(face = "bold", hjust = 0.5),
    legend.position = "none"
  ) +
  labs(
    title = "Shannon Entropy (Lipid Diversity) Across Treatments",
    x = "Treatment",
    y = "Shannon Entropy (H_j)"
  ) +
  annotate("text", Inf, Inf, 
           label = paste0("ANOVA p-value = ", round(anova_pvalue, 3)),
           hjust = 1.1, vjust = 1.5, size = 5, fontface = "bold", color = "black")

# Display plot
print(shannon_plot)

# Save plot
ggsave("shannon_entropy_anova_plot.png", plot = shannon_plot, width = 10, height = 7, dpi = 300)

```
